\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Journal: Ecosphere. Manuscript ID: ECS23-0342. Manuscript type: Article}}

### 2023-10-30 manuscript/manuscriptRW.Rmd
small correction in the manuscript, last version before diff

title: "\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Submitted to Ecosphere. Manuscript type: Article}}"

@@

- Christopher P.O. Reyer$^8$,

@@

       $^6$Institute of Silviculture, Department of Forest and Soil Sciences, University of Natural Resources and Life Sciences, Peter Jordan Strasse 82, A-1190 Vienna, Austria \newline

@@

       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen, Slovakia \newline

@@

Each of the models reported the number of trees crossing the two
diameter thresholds by sampling 200 times in a 10-year interval for each
species and per ha for each of the 200 sites. Multiple samples per site
were used to better understand the simulated variation within each site.
This was done using different strategies, depending on the model,
including (1) sampling simulated data from the same 1 ha plot in the
equilibrium over time, (2) sampling several 1 ha plots from the
simulated forest at one specific point in time (in the equilibrium) or
(3) a combination of (1) and (2). This resulted in 880,000 observations
per model that simulated the 11 species included in the protocol (200
sites, 200 samples per site, 11 species and two diameter thresholds).
For the models that simulated additional species, their regeneration
rates were aggregated as "others".

@@

```{=tex}
\begin{equation}
Reineke = \frac{N7}{N10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on
(1) model complexity as defined by @bugmann2022, (2) model type
(empirical or process-based), (3) the presence or absence of a canopy
feedback for regeneration, and (4) the scale of application of the model
(stand, landscape or global). We tested for significant differences
between simulated and observed values of the recruitment levels and
species diversity by using paired t-tests.

@@

O. Díaz-Yáñez and Y. Käber contributed equally and share the first
authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
design. Y. Käber and O. Díaz-Yáñez prepared the input data used by all
modeling teams. All authors contributed by conducting the simulations
with their respective models. O. Díaz-Yáñez led the analysis of the
simulation results with contributions during the workshop from Y.Käber,
T. Anders, K. H. Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T.
Hickler, H. Lischke, M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X.
Morin, W. Rammer, and D. Scherrer. O. Díaz-Yáñez and H. Bugmann led the
writing of the manuscript and provided the first draft. All authors
participated in the revision of the manuscript and approved its
submission.

@@

@@

```{r tableModels, echo=FALSE,  message=FALSE, warning=TRUE}
library(knitr)
modelDF <- readr::read_csv(here::here("manuscript", "modelsDescription.csv"))
modelDF$Sampling <- NULL

@@

kable(modelDF,
      format = "latex", booktabs = TRUE,
      caption = "") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |>
      kableExtra::column_spec(column = 2, width = "3em") |>
      kableExtra::column_spec(column = 3, width = "4em") |>
      kableExtra::column_spec(column = 4, width = "5em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "6em") |>
      kableExtra::column_spec(column = 7, width = "5em") |>
      kableExtra::column_spec(column = 8, width = "8em") |>
      kableExtra::column_spec(column = 9, width = "8em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "5em") |>
      kableExtra::column_spec(column = 12, width = "10em") |>
      kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
                                 "PB-Process based, EM-Empirical",
                                 "C-Cohort, I-Individual",
                                 "Y-Yes, N-No",
                                 "R-Regeneration, Rc-Recruitment",
                                 "D-Daily, M-Monthly"),

@@

# Table and Figure captions

**Table 1.** Models included in this study, their characteristics and
simulation strategies used. *Scale* refers to the models scale, *Type*
to the models type, *Population structure* to the model population
approach, *Feedback* to the inclusion or not of adult tree feedback to
the level and species composition of regeneration, *Approach* to the
model regeneration module approach, *Start from* to the stage from which
trees are recruited, *Complexity*, to the mean regeneration formulation
as estimated in (Bugmann & Seidl, 2022), *Species* to which species from
the original eleven species were simulated, *Runtime* (spin up) [years]
to the number of years used per model in their simulations for the spin
up period and regeneration sampling; *Climate data type (use of years)*
to the type of data used in the simulations and how it was used, and
*Reference* provides the link to the publication with more model
details. Further information can be found on the individual model
reports available in the supplementary materials.

>>>>

title: "\\fontsize{12pt}{3pt}\\selectfont \\textbf{\\textit{Journal: Ecosphere. Manuscript ID: ECS23-0342. Manuscript type: Article}}"

@@

- Christopher P.O. Reyer$^{15}$,

@@

       $^6$Institute of Silviculture, Department of Forest and Soil Sciences, University of Natural Resources and Life Sciences, Peter Jordan Strasse 82, A-1190 Vienna, Austria. \newline

@@

       $^{10}$Department of Forest Harvesting, Logistics and Ameliorations, Faculty of Forestry, Technical University in Zvolen, Slovakia. \newline

@@

Each of the models reported the number of new trees crossing the two
diameter thresholds by sampling 200 times in a 10-year interval for each
species and per ha for each of the 200 sites. Multiple samples per site
were used to better understand the simulated variation within each site.
This was done using different strategies, depending on the model,
including (1) sampling simulated data from the same 1 ha plot in the
equilibrium over time, (2) sampling several 1 ha plots from the
simulated forest at one specific point in time (in the equilibrium) or
(3) a combination of (1) and (2). This resulted in 880,000 observations
per model that simulated the 11 species included in the protocol (200
sites, 200 samples per site, 11 species and two diameter thresholds).
For the models that simulated additional species, their regeneration
rates were aggregated as "others".

@@

```{=tex}
\begin{equation}
Reineke = \frac{N7}{N10}^{-1.605} \label{eqReineke}
\end{equation}
```
We assessed model performance in relation to model traits focusing on
(1) model complexity as defined by @bugmann2022, (2) model type
(empirical or process-based), (3) the presence or absence of a canopy
feedback for regeneration, and (4) the scale of application of the model
(stand, landscape or global). We tested for significant differences
between simulated and observed values of the recruitment levels and
species diversity by using t-tests with a correction for multiple
testing. Note that the t-tests does not present conclusive evidence for
a specific hypotheses, but rather facilitates managing the extensive
number of comparisons discussed in our study.

@@

@@

O. Díaz-Yáñez and Y. Käber contributed equally and share the first
authorship. H. Bugmann, Y. Käber, and O. Díaz-Yáñez developed the study
design. Y. Käber and O. Díaz-Yáñez prepared the input data used by all
modeling teams. All authors contributed by conducting the simulations
with their respective models. O. Díaz-Yáñez led the analysis of the
simulation results with contributions during the workshop from Y. Käber,
T. Anders, K. H. Braziunas, J. Brůna, S. M. Fischer, J. Hetzer, T.
Hickler, H. Lischke, M.Mahnken, P. Mairota, K. Merganičová, T. Mette, X.
Morin, W. Rammer, and D. Scherrer. O. Díaz-Yáñez and H. Bugmann led the
writing of the manuscript and provided the first draft. All authors
participated in the revision of the manuscript and approved its
submission.

@@

**Table 1.** Models included in this study, their characteristics and
simulation strategies used. Scale refers to the models scale, Type to
the models type, Population structure to the model population approach,
Feedback to the inclusion or not of adult tree feedback to the level and
species composition of regeneration, Approach to the model regeneration
module approach, Start from to the stage from which trees are recruited,
Complexity, to the mean regeneration formulation as estimated in
(Bugmann & Seidl, 2022), Species to which species from the original
eleven species were simulated, Runtime (spin up) [years] to the number
of years used per model in their simulations for the spin up period and
regeneration sampling; Climate data type (use of years) to the type of
data used in the simulations and how it was used, and Reference provides
the link to the publication with more model details. Further information
can be found on the individual model reports available in the
supplementary materials.

```{r tableModels, echo=FALSE,  message=FALSE, warning=TRUE}
library(knitr)
modelDF <- readr::read_csv(here::here("manuscript", "modelsDescription.csv"))
modelDF$Sampling <- NULL
captionText <- ""

@@

kable(modelDF,
      format = "latex", booktabs = TRUE,
      caption = captionText) |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |>
      kableExtra::column_spec(column = 2, width = "3em") |>
      kableExtra::column_spec(column = 3, width = "4em") |>
      kableExtra::column_spec(column = 4, width = "5em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "6em") |>
      kableExtra::column_spec(column = 7, width = "5em") |>
      kableExtra::column_spec(column = 8, width = "8em") |>
      kableExtra::column_spec(column = 9, width = "8em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "5em") |>
      kableExtra::column_spec(column = 12, width = "10em") |>
      kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
                                 "PB-Process based, EM-Empirical",
                                 "C-Cohort, I-Individual",
                                 "Y-Yes, N-No",
                                 "R-Regeneration, Rc-Recruitment",
                                 "D-Daily, M-Monthly"),

@@

# Figure captions
